---
title: "nativeORT"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{nativeORT}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nativeORT)

model_path <- file.path(tools::R_user_dir("nativeORT", "data"), "yolo11n.onnx")

if (!file.exists(model_path)) {
  message("Downloading YOLOv11 model (~12Mb)...")
  dir.create(dirname(model_path), recursive=TRUE, showWarnings = FALSE)
  utils::download.file(
    url="https://github.com/ultralytics/assets/releases/download/v8.4.0/yolo11n.onnx",
    destfile=model_path,
    mode="wb"
  )
  message("yolo11n.onnx downloaded!")
}
```

## Description

nativeORT is the only R-native ONNX inference engine with no Python or reticulate dependencies. Built on native C++ bindings, nativeORT can run at real-time speeds to edge deployments.

nativeORT is easy to use and works with any .onnx file supported by onnxruntime v1.25.0.

## How to Use

### onnxruntime Session Creation

First, create an ORT session with any old .onnx file

```{r}
session <- nativeORT::ort_session(model_path)
print(session)
```

### Inference

Now, we can just pass any input and run inference

```{r}
# typical RGB 256x256 image
input <- array(
  runif(1 * 3 * 256 * 256),
  dim=c(1L, 3L, 256L, 256L)
)

raw <- nativeORT::ort_infer_raw(session, input)
dim(raw)
```

and that's it! this works for any ONNX model you may want run!

### Timing

nativeORT is fast! The Benchmark vignette gives more detail, but just for example, let's run a quick timing test here:

```{r}
dir.create(path.expand("~/.nativeORT/cache"),
           recursive = TRUE, showWarnings = FALSE
           )
session <- nativeORT::ort_session(model_path,
                                  provider='coreml',
                                  cache_dir=path.expand("~/.nativeORT/cache")
           )
# warmup
message("warming up CoreML")
for (i in 1:10) warmup <- nativeORT::ort_infer_raw(session, input)

t_inference <- system.time(nativeORT::ort_infer_raw(session, input))
message(sprintf("ort_infer_raw: %.1f ms", t_inference["elapsed"] * 1000))
```

